---
title: "QC_Automation"
author: "Brian Yandell"
date: "2024-09-09"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(metabr)
```

This is based on new DO data from November 2025.
Assume we only have raw data, not hand-corrected.

# QC steps and normalization (Batch Correction)

Need to correct the two modes (`HC_Neg` and `RP_Pos`) separately.
`HC_Neg` mode targeted data of founder samples used as example
to showed how Qiushi did batch correction manually.
See Batch_Correction_Procedure_RC2_Project.pdf
(file `data/QC Batch Correction/Batch_Correction_Procedure_RC2_Project.pdf`)

```{r}
# Raw data in CSV
dirpath <- 'data/2025_Nov_Raw'
```

## Individual QC Steps

Shown for `HC_Neg` data.
For `RP_Pos` data, exclude every 1st QC run on plate as it
historically was always an outlier.
QC steps were verified against hand-calculations.
See [QC_steps.Rmd](https://github.com/byandell/metabr/blob/master/QC_steps.Rmd)
section titles `Examine QC data`.

1. Transpose Data and Extract Info from Sample Names

Note addition of `BatchPlate_HCneg` as QC is by `Plate` and `Batch`, not `rundate`.

```{r}
raw_HCneg <- read_raw_metab(dirpath,
  "DO_Livers_HC_Neg_Targeted(in).csv")
#  Batch_Plate = BatchPlate_HCneg)
```

Fake Batch and Plate

```{r}
BatchPlate_HCneg <- dplyr::distinct(raw_HCneg, rundate) |>
  dplyr::mutate(Batch = 1, Plate = dplyr::row_number())
raw_HCneg <- dplyr::left_join(
  raw_HCneg,
  BatchPlate_HCneg,
  by = "rundate")
```

2. Only pick up QC runs of every plate, and calculate correction factors (CF) for each metabolite.
    + Average of QC per plate/Average of QC all plate.

```{r}
qc_HCneg <- calc_cf(raw_HCneg)
```

3. If metabolites weren’t measured in (**all?**) QC samples, but were measured in study samples: use the CF average of all metabolites in one plate as the CF of the missing metabolites on that plate.

```{r}
qc_HCneg <- replace_missing_ave_cf(qc_HCneg)
```

4. Correct the Peak Area using CF for each metabolite 

```{r}
rescale_HCneg <- correct_raw_cf(raw_HCneg, qc_HCneg)
```

5. Write out file

```{r}
wrpath <- "data/2025_Nov_Automated"
if(!dir.exists(wrpath)) dir.create(wrpath)
write_metab(rescale_HCneg,
            file.path(wrpath, "HC_Neg_Targeted.csv"))
```

## Combine QC Steps Into One Routine.

```{r}
rescale_HCneg <- qc_steps(handpath,
  "Founder_Plasma_HC_Neg_Targeted_Batch_Corrected_Data.xlsx",
  FALSE, BatchPlate_HCneg)
```

```{r}
rescale_RPpos <- qc_steps(handpath,
  "Founder_Plasma_RP_Pos_targeted_Batch_Corrected_Data.xlsx",
  TRUE, BatchPlate_RPpos)
```

# Untargeted Data

```{r}
rescale_HCnegU <- qc_steps(handpath,
  "Founder_Plasma_HC_Neg_untargeted_All_Batch_Corrected_Data.xlsx",
  FALSE, BatchPlate_HCneg)
```

```{r}
rescale_RPposU <- qc_steps(handpath,
  "Founder_Plasma_RP_pos_untargeted_All_Batch_Corrected_Data.xlsx",
  TRUE, BatchPlate_RPpos)
```

# Write Out Files

```{r}
wrpath <- "data/2025_Nov_Automated"
if(!dir.exists(wrpath)) dir.create(wrpath)
write_metab(rescale_HCneg,
            file.path(wrpath, "HC_Neg_Targeted.csv"))
write_metab(rescale_HCnegU,
            file.path(wrpath, "HC_Neg_Untargeted.csv"), FALSE)
write_metab(rescale_RPpos,
            file.path(wrpath, "RP_Pos_Targeted.csv"))
write_metab(rescale_RPposU, 
            file.path(wrpath, "RP_Pos_Untargeted.csv"), FALSE)
```

```{r}
HCneg_Tar <- rescale_HCneg |>
  dplyr::select(
  sample, run, rundate, Batch, Plate, mouse_id, time, rep, compound, value) |>
  tidyr::pivot_wider(id_cols = sample:rep,
    names_from = "compound", values_from = "value")
```

```{r}
HCneg_UnTar <- rescale_HCnegU |>
  dplyr::select(
  sample, run, rundate, Batch, Plate, mouse_id, time, rep, compound, medRt, value) |>
  dplyr::mutate(medRt = round(medRt, 3)) |>
  tidyr::unite(compound, compound, medRt) |>
  tidyr::pivot_wider(id_cols = sample:rep,
    names_from = "compound", values_from = "value")
```

